---
title: "Relationship between nutritional content and recipe ratings"
author: "Evelin Haasz"
format: html
---
```{r setup}
library(tidyverse)
```

## Introduction
Online recipe platforms collect and provide various different types of data about recipes and their different parameters. One commonly used indicator of recipe quality is the rating of users. At the same time there might be other parameters, such as nutritional values, that can influence how appealing a recipe is perceived.

The aim of this project is to exmaine whether there is an association between the caloric content of recipes and their average user ratings using the data from the TidyTuesday package. Specifically, the hypothesis that recipes with higher calories recieive higher agerage user ratings.

## Data

The data used for analyses comes from the TidyTuesday dataset published on 2025-09-16, which containes information about recipes sourced from Allrecipes.

```{r}
all_recipes_o <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-09-16/all_recipes.csv")
```

In the analyses we will focus on the following variables:
- *avg_rating*: Average user rating of a recipe
- *calories*: Caloric content of the recipe
- *total_time*: Total preparation time
- *protein, fat, carbs*: Macronutrient information

```{r}
all_recipes_c <- all_recipes_o %>%
select(name, avg_rating, calories, total_time, protein, fat, carbs) %>%
drop_na()
head(all_recipes_c)
summary(all_recipes_c)
```

## Exploratory data analysis
